Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone

### 2020-11-19 paper/ifo-conclusions.tex
full read-through before submission

In this paper, we use a table-top Michelson interferometer as an analogue to a gravitational-wave detector, demonstrating a range of signal processing techniques used within the gravitational-wave community and more broadly. 
Gravitational-wave signals are simulated by playing sound recordings through a speaker fixed to the back of one of the arm mirrors. 
We consider quasi-monochromatic signals as an analogue to continuous wave signals to directly demonstrate techniques used in current continuous gravitational-wave searches.
We then consider more complex audio (music and speech) for a more general interest in signal processing and explore use of the interferometer as an optical microphone. 
A variety of example activities are presented. 
We find that the logMMSE produces the best results, however it does not enhance the recording to an intelligible level. 
The complex audio signals can also serve as more distant analogues to un-modelled gravitaitonal-wave burst signals, e.g. from supernova.

\begin{comment}
% included these paras in the first to avoid repetition and reduce some of the length of the paper before submission
% reduced detail for conclusion (trying to reduce overall length)
In Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} we analyse audio signals with constant and wandering frequencies respectively as analogues to continuous gravitational wave signals. 
The wandering frequency signal is recovered using using the Viterbi algorithm. 
These sections directly demonstrate the techniques used in current continuous gravitational-wave searches. 
%We demonstrate the recovery of a signal with constant frequency and signal with slowly wandering frequency in Sections~\ref{sec:ifo} and ~\ref{sec:single_tone}, respectively. We use a basic webcam with a sampling rate of $30\,{\rm Hz}$ to track the interference pattern. For the wandering frequency we successfully recover the injected signal using the Viterbi algorithm. These sections directly demonstrate the techniques used in continuous gravitational-wave searches.

@@

The demonstration presented has potential to be adapted for use in the undergraduate laboratory.
It can be used to teach topics of interest to physics and electrical engineering students including, interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques. 
The increased excitement and public interest in the field of gravitational-wave research in recent years means that this demonstration may also be adapted as a tool for explaining gravitational-wave research to a wider, non-specialist audience. 
The field of gravitational-wave astrophysics continuous to grow and the future will bring many more detections of binary black holes and neutron stars, as well as the anticipated first detection of other classes of signals such as continuous waves to which this demonstration provides some charming insights. 
%The demonstration presented here has great potential to be adapted for use in the undergraduate laboratory. It can be used to demonstrate topics such as interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques to physics and electrical engineering students. Given the increased level of excitement and public interest in the field of gravitational-wave astronomy in recent years, this table-top demonstration can be used as a tool for explaining gravitational-wave research to a wider, non-specialist audience. The future will bring many more detections, including further observations of transient signals from binary black hole and binary neutron star mergers, as well as the anticipated first detections of other classes of signal, including continuous waves, to which this demonstration provides some charming insights.

>>>>

In this paper, we use a table-top Michelson interferometer as an analogue to a gravitational wave detector, demonstrating a range of signal processing techniques used within the gravitational wave community and more broadly. 
Gravitational wave signals are simulated by playing sound recordings through a speaker fixed to the back of one of the arm mirrors. 
We consider quasi-monochromatic signals as an analogue to continuous wave signals to directly demonstrate techniques used in current continuous gravitational wave searches.
We then consider more complex audio (music and speech) for a more general interest in signal processing and explore use of the interferometer as an optical microphone. 
A variety of example activities are presented. 
We find that the logMMSE produces the best results, however it does not enhance the recording to an intelligible level. 
The complex audio signals can also serve as more distant analogues to un-modelled gravitaitonal-wave burst signals, e.g. from supernova.

\begin{comment}
% included these paras in the first to avoid repetition and reduce some of the length of the paper before submission
% reduced detail for conclusion (trying to reduce overall length)
In Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} we analyse audio signals with constant and wandering frequencies respectively as analogues to continuous gravitational wave signals. 
The wandering frequency signal is recovered using using the Viterbi algorithm. 
These sections directly demonstrate the techniques used in current continuous gravitational wave searches. 
%We demonstrate the recovery of a signal with constant frequency and signal with slowly wandering frequency in Sections~\ref{sec:ifo} and ~\ref{sec:single_tone}, respectively. We use a basic webcam with a sampling rate of $30\,{\rm Hz}$ to track the interference pattern. For the wandering frequency we successfully recover the injected signal using the Viterbi algorithm. These sections directly demonstrate the techniques used in continuous gravitational wave searches.

@@

The demonstration presented has potential to be adapted for use in the undergraduate laboratory.
It can be used to teach topics of interest to physics and electrical engineering students including, interferometer physics; gravitational wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques. 
The increased excitement and public interest in the field of gravitational wave research in recent years means that this demonstration may also be adapted as a tool for explaining gravitational wave research to a wider, non-specialist audience. 
The field of gravitational wave astrophysics continuous to grow and the future will bring many more detections of binary black holes and neutron stars, as well as the anticipated first detection of other classes of signals, such as continuous waves, to which this demonstration provides some charming insights. 
% \jam{``charming'' is very sweet}
%The demonstration presented here has great potential to be adapted for use in the undergraduate laboratory. It can be used to demonstrate topics such as interferometer physics; gravitational wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques to physics and electrical engineering students. Given the increased level of excitement and public interest in the field of gravitational wave astronomy in recent years, this table-top demonstration can be used as a tool for explaining gravitational wave research to a wider, non-specialist audience. The future will bring many more detections, including further observations of transient signals from binary black hole and binary neutron star mergers, as well as the anticipated first detections of other classes of signal, including continuous waves, to which this demonstration provides some charming insights.

### 2020-11-19 paper/ifo-futureWork.tex
full read-through before submission

The optical microphone experiment described here can be used to demonstrate and teach a variety of topics including basic optics and photodiode circuits, to signals processing and speech enhancement. Here we consider further extensions and improvements to the experiment such as noise isolation, transfer function analysis, speech recognition, and control systems.

% reduced length
Isolating the experiment from interference may provide some reduction in noise. 
For example, the Raspberry Pi could be powered with a commercial battery instead of mains power. 
%An initial improvement could be to better isolate the experiment, especially from the sources of mains noise. 
%As the Raspberry Pi demands little power and connects to USB it could instead be powered with a commercial battery rather than the mains supply. 
All other components of the circuit are connected to the Pi. 
%All other components of the circuit, including the photodiode are connected to the Pi. 
The laser is DC powered through a converter, so should not couple noise from the mains into the interferometer. 
Transferring the circuit from a breadboard to a printed circuit board can also reduce interference~\cite{elfekey2013design}.
%With the interferometer in a dark room and under a box to better isolate from any ambient light or air currents, this would hopefully remove (or better separate) all sources of the 50Hz mains hum from the optical microphone.
A feedback loop control system can also be used to suppress unwanted movement of the interferometer components~\citep{abbott2017exploring, Sekiguchi:2016bmv, verhoeven2009robust}, similar to methods used to isolate gravitational wave observatories. 
%We can consider improving the stability of the optical microphone components. A feedback loop control system can be used to suppress any unwanted movement of the interferometer components~\citep{abbott2017exploring, Sekiguchi:2016bmv, verhoeven2009robust}.

@@

A thorough characterisation of the systems transfer function would allow us to understand how audio signals couple through the interferometer. 
%We can also benefit from fully characterizing the transfer function of the system as it provides the exact information of how audio signal couples through the interferometer. 
The total transfer function starts with the voltage being sent to the speaker and ends with the voltage recorded by the Pi. It also would include any inherent non-linearities in the speaker, the speaker-mirror coupling, the path length to intensity relation (Appendix~\ref{app:intensity_derivation}), and the photodiode. The schematic in Fig.~\ref{fig:pipeline_highlighted} shows the signal flow through the system.
% commented out whilst James is checking this
% jam: I'm pretty confident this is what Rob said. 
%With perfect knowledge of the transfer function, signal recovery is limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
%\jam{[check with Rob et al. about the veracity of this]}

A voluminous literature exists on using hidden Markov models trained on phonemes to recognize speech~\cite{HMM_english}. These could perform the final stage of speech enhancement for the optical microphone. Interestingly, this connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a hidden Markov model, albeit of a different type. Alternatively, machine learning solutions exist throughout the field that are able to compete with statistical techniques such as the logMMSE estimator~\cite{SEGAN}.

We could also use more traditional techniques such as wavelet transform (WT) \citep{nason1995stationary} to extract the signal from noise, and compare these with the above methods. WT provides both time and frequency information, making it easier to pinpoint the origin of noise with respect to time. In Refs.~\cite{tufekci2000feature,agbinya1996discrete} WT methods are proposed for speech recognition.

Returning to the demonstration of continuous gravitational wave analysis, further additions could include analysing data from multiple interferometers to extract directional information for audio signals. 
%Finally, considering the desire to demonstrate continuous wave analysis, further improvements to the optical microphone could lead to more advanced demonstration of techniques such as aiming or locking two or more optical microphones in an array.
% I think locking is the wrong word to use here as it is a word used a lot in GW instrumentation (an interferometer can be 'locked' on its own). I've tried to rephrase based on what I think you are getting at. -hannah

>>>>

The optical microphone experiment described here can be used to demonstrate and teach a variety of topics, including basic optics and photodiode circuits, to signals processing and speech enhancement. Here, we consider further extensions and improvements to the experiment such as noise isolation, transfer function analysis, speech recognition, and control systems.

% reduced length
Isolating the experiment from interference may provide some reduction in noise. 
For example, the Raspberry Pi could be powered with a commercial battery instead of mains power. 
%An initial improvement could be to better isolate the experiment, especially from the sources of mains noise. 
%As the Raspberry Pi demands little power and connects to USB it could instead be powered with a commercial battery rather than the mains supply. 
All other components of the circuit draw power from the Pi, so this would isolate the circuit from the mains. 
%All other components of the circuit, including the photodiode are connected to the Pi. 
The laser is DC powered through a converter, so should not couple noise from the mains into the interferometer. 
Transferring the circuit from a breadboard to a printed circuit board can also reduce interference~\cite{elfekey2013design}.
%With the interferometer in a dark room and under a box to better isolate from any ambient light or air currents, this would hopefully remove (or better separate) all sources of the 50Hz mains hum from the optical microphone.
A feedback loop control system can also be used to suppress unwanted movement of the interferometer components~\citep{abbott2017exploring, Sekiguchi:2016bmv, verhoeven2009robust}, similar to methods used to isolate gravitational wave observatories. 
%We can consider improving the stability of the optical microphone components. A feedback loop control system can be used to suppress any unwanted movement of the interferometer components~\citep{abbott2017exploring, Sekiguchi:2016bmv, verhoeven2009robust}.

@@

A thorough characterisation of the systems transfer function would allow us to understand how audio signals couple through the interferometer. 
%We can also benefit from fully characterizing the transfer function of the system as it provides the exact information of how audio signal couples through the interferometer. 
The total transfer function starts with the voltage being sent to the speaker and ends with the voltage recorded by the Pi. It also would include any inherent non-linearities in the speaker, the speaker-mirror coupling, the path length to intensity relation (Appendix~\ref{app:intensity_derivation}), and the photodiode. The schematic in Fig.~\ref{fig:pipeline_highlighted} shows the abstract signal flow through the system.
% commented out whilst James is checking this
% jam: I'm pretty confident this is what Rob said. 
%With perfect knowledge of the transfer function, signal recovery is limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
%\jam{[check with Rob et al. about the veracity of this]}

A voluminous literature exists on using hidden Markov models trained on phonemes to recognize speech~\cite{HMM_english}. These could perform the final stage of speech enhancement for the optical microphone. Interestingly, this coincidently connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a hidden Markov model, albeit of a different type. Alternatively, machine learning solutions exist throughout the field that are able to compete with statistical techniques such as the logMMSE estimator~\cite{SEGAN}.

We could also use more traditional techniques such as wavelet transform (WT) \citep{nason1995stationary} to extract the signal from noise, and compare these with the above methods. WT provides both time and frequency information, making it easier to pinpoint the origin of noise with respect to time. In Refs.~\cite{tufekci2000feature,agbinya1996discrete}, WT methods are proposed for speech recognition.

Returning to the demonstration of continuous gravitational wave analysis, further extensions could include analysing data from multiple interferometers to extract directional information for audio signals. Although, this would require coupling from a separate noise source instead of from a speaker attached directly to one of the mirrors of the interferometer. 
%Finally, considering the desire to demonstrate continuous wave analysis, further improvements to the optical microphone could lead to more advanced demonstration of techniques such as aiming or locking two or more optical microphones in an array.
% I think locking is the wrong word to use here as it is a word used a lot in GW instrumentation (an interferometer can be 'locked' on its own). I've tried to rephrase based on what I think you are getting at. -hannah

### 2020-11-19 paper/ifo-introduction.tex
full read-through before submission

In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics and revealed a new means to observe the universe~\cite{GW150914}.
% What are gws and gw detectors 
Gravitational waves are a prediction of Albert Einstein's theory of general relativity~\cite{Einstein:1916}, 
They are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
The effect of gravitational waves is a change in lengths: a `stretching and squeezing' of the distance between objects. 
Ground-based gravitational-wave observatories such as the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} use the interference of laser light to measure changes in distance. 
Gravitational-wave observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
Table-top interferometers are commonly used in undergraduate laboratory experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,AMIGO:online,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.

% CW 
To date the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020} and binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020}, as well as a black hole merger with an object that could either have been a neutron star or another black hole~\cite{GW190814}. 
However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
Continuous waves may be emitted at a constant frequency or may wander slowly in frequency over time. 
Rotating neutron stars are prime candidates for continuous wave emission, especially those in low mass X-ray binaries (LMXB), where the neutron star is in orbit with a low mass stellar companion.
%where a compact object such as a black hole or a neutron star is in orbit with a low mass stellar companion.
% we're talking about neutron stars in LMXB's, so we don't need to mention black holes in LMXB's.
The rotation frequency of the neutron star can wander over time due to variable accretion of matter (and hence angular momentum transfer) from the stellar companion~\cite{xraybinaries:1997}. 
Scorpius X-1 is a prime LMXB target for continuous wave searches. 
%Scorpius X-1 is an LMXB which is particularly bright in X-rays, making it a prime target for continuous wave searches. %from LMXBs. 
Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, RadiometerO1O2:2019, SearchRadiometerO1:2017, SearchCrossCorrO1:2017, ScoX1ViterbiO1:2017, SearchTwoSpecS6:2017, MiddletonEtAlO2LMXBs:2020}.

% changed present to use, as this is not the first time a Michelson has been presented as a tpy GW detector - hannah
In this paper, we use a table-top Michelson interferometer as a toy gravitational-wave detector designed to detect sound instead of gravitational waves. 
We then extend its use to an \emph{optical microphone} (using light to capture sound) and present a range of example analysis techniques for educators to use. 
% re-added "as an undergraduate experiment to be clear we are not talking about school teaching - hannah
As an undergraduate lab experiment, the demonstration can be used to teach topics ranging from continuous-wave detection and analysis to electronics, signal processing, and speech enhancement. 
%% I don't like the use of ``freedom to explore''. 
% removed Additionally - I don't think it is neccessary - hannah
It allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, opto-mechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019, SuvorovaEtAl:2017,SuvorovaEtAl:2016}. 
% This demonstration can be adapted to an effective outreach tool to explain gravitational wave science to a non-specialist audience.
% Rephrased as this kind of sounds like we are an exhibition too [hannah] - This demonstration also has the potential to be used as an outreach tool alongside other exhibitions
%This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities including; hands-on demonstrations and exhibitions~\cite{TTExhibit:2020,CavagliaExhibit:2009,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational-wave research groups around the world. 
This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational-wave research groups around the world~\cite{TTExhibit:2020,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online,chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}. 
These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational-wave science to non-specialist audiences.
%\han{look up Sarah Farmer's project}

This paper is laid out as follows. 
In Section~\ref{sec:ifo} we detail the table-top interferometer design. 
We demonstrate observing a single tone from a speaker in Section~\ref{sec:single_tone}. 
In Section~\ref{sec:viterbi_wandering} we replace the single tone with a wandering frequency signal and analyse the signal using a hidden Markov model technique (the Viterbi algorithm) from continuous wave searches. 
In Section~\ref{sec:optical_microphone}, we demonstrate capture and playback of complex audio, such as music and speech.
% removed as there's not space to really say what this means here: to varying degrees of success.
This demonstration of an optical microphone  serves as a more general exhibition of signal processing with a range of examples which can be used in the undergraduate laboratory. 
We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in ~\ref{sec:conclusions}. 
Appendix~\ref{app:code} collects the software and scripts used to produce this work.

>>>>

In 2015, the first detection of gravitational waves from the merger of two black holes~\cite{GW150914} marked a breakthrough in modern astrophysics and revealed a new means to observe the universe.
% What are gws and gw detectors 
Gravitational waves are a prediction of Albert Einstein's theory of general relativity~\cite{Einstein:1916}, 
They are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
The effect of gravitational waves is a change in lengths: a `stretching and squashing' of the distance between objects. 
Ground-based gravitational wave observatories such as the Advanced Laser Interferometer Gravitational wave Observatory (LIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} use the interference of laser light to measure changes in distance. 
Gravitational wave observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
Table-top interferometers are commonly used in undergraduate laboratory experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,AMIGO:online,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.

% CW 
To date the network of gravitational wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020} and binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020}, as well as a black hole merger with an object that could either have been a neutron star or another black hole~\cite{GW190814}. 
However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
Continuous waves may be emitted at a constant frequency or may wander slowly in frequency over time. 
Rotating neutron stars are prime candidates for continuous wave emission, especially those in low mass X-ray binaries (LMXB), where the neutron star is in orbit with a low mass stellar companion.
%where a compact object such as a black hole or a neutron star is in orbit with a low mass stellar companion.
% we're talking about neutron stars in LMXB's, so we don't need to mention black holes in LMXB's.
The rotation frequency of the neutron star can wander over time due to variable accretion of matter (and hence angular momentum transfer) from the stellar companion~\cite{xraybinaries:1997}. 
Scorpius X-1 is a prime LMXB target for continuous wave searches. 
%Scorpius X-1 is an LMXB which is particularly bright in X-rays, making it a prime target for continuous wave searches. %from LMXBs. 
Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, RadiometerO1O2:2019, SearchRadiometerO1:2017, SearchCrossCorrO1:2017, ScoX1ViterbiO1:2017, SearchTwoSpecS6:2017, MiddletonEtAlO2LMXBs:2020}.

% changed present to use, as this is not the first time a Michelson has been presented as a tpy GW detector - hannah
In this paper, we use a table-top Michelson interferometer as a toy gravitational wave detector designed to detect sound instead of gravitational waves. 
We then extend its use to an ``optical microphone'' (using light to capture sound) and present a range of example analysis techniques for educators to use. 
% re-added "as an undergraduate experiment to be clear we are not talking about school teaching - hannah
As an undergraduate lab experiment, the demonstration can be used to teach topics ranging from continuous-wave detection and analysis to electronics, signal processing, and speech enhancement. 
%% I don't like the use of ``freedom to explore''. 
% removed Additionally - I don't think it is neccessary - hannah
It allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, opto-mechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019, SuvorovaEtAl:2017,SuvorovaEtAl:2016}. 
% This demonstration can be adapted to an effective outreach tool to explain gravitational wave science to a non-specialist audience.
% Rephrased as this kind of sounds like we are an exhibition too [hannah] - This demonstration also has the potential to be used as an outreach tool alongside other exhibitions
%This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities including; hands-on demonstrations and exhibitions~\cite{TTExhibit:2020,CavagliaExhibit:2009,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational wave research groups around the world. 
This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational wave research groups around the world~\cite{TTExhibit:2020,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online,chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}. 
These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational wave science to non-specialist audiences.
%\han{look up Sarah Farmer's project}

This paper is laid out as follows. 
In Section~\ref{sec:ifo}, we detail the table-top interferometer design. 
In Section~\ref{sec:single_tone}, we demonstrate observing a single tone from a speaker. 
In Section~\ref{sec:viterbi_wandering}, we replace the single tone with a wandering frequency signal and analyse the signal using a hidden Markov model technique (the Viterbi algorithm) from continuous wave searches. 
In Section~\ref{sec:optical_microphone}, we demonstrate capture and playback of complex audio, such as music and speech.
% removed as there's not space to really say what this means here: to varying degrees of success.
This demonstration of an optical microphone  serves as a more general exhibition of signal processing with a range of examples which can be used in the undergraduate laboratory. 
We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}. 
We collect the software and scripts used to produce this work in Appendix~\ref{app:code}.

### 2020-11-19 paper/ifo-singleTone.tex
combined notch and Wiener plots
full read-through before submission

\begin{figure}
	\includegraphics[width=.49\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
	\caption{\label{fig:webcam_spectrum}
Recovery of a tone at a constant frequency: Here we plot the Fourier amplitude of the intensity pattern against frequency.
The injected signal has a frequency of $2.09\,{\rm Hz}$, while the recovered signal peaks at $2.099\,{\rm Hz}$ and has a FWHM of $0.033\,{\rm Hz}$.
The plot also shows two harmonics at $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ with normalised amplitude to the primary peak of $0.103$ and $0.098$ respectively.
}	
\end{figure}

@@

Continuous-wave searches look for nearly monochromatic signals. In this section we consider a simple sinusoidal tone; a single constant frequency.

As described in Section~\ref{sec:ifo}, the audio signal is played through a speaker fixed to the back of M2. 
The intensity of the interference pattern is measured at a single point on the screen, indicated by the pink star in Fig~\ref{fig:interference_pattern}. 
The webcam records video in three colour channels; red, blue, and green. 
We take the green channel as an approximation of the total intensity produced by the monochromatic green laser.
The webcam samples at a rate of $30\, {\rm Hz}$, which limits the spectral content of observable signals to less than $15\,{\rm Hz}$, the Nyquist frequency.

>>>>

\begin{figure}
	\includegraphics[width=.49\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
	\caption{\label{fig:webcam_spectrum}
Recovery of a tone at a constant frequency: Here we plot the Fourier amplitude of the intensity pattern against frequency.
The injected signal has a frequency of $2.09\,{\rm Hz}$, while the recovered signal peaks at $2.099\,{\rm Hz}$ and has a FWHM of $0.033\,{\rm Hz}$.
The plot also shows two harmonics at $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ with amplitudes $10.3 \%$ and $9.8 \%$ of the amplitude of the primary peak, respectively.
}	
\end{figure}

@@

Continuous-wave searches look for nearly monochromatic signals. In this section we consider a simple sinusoidal tone at a single, constant frequency. %should be single constant frequency *signal*?

As described in Section~\ref{sec:ifo}, the audio signal is played through a speaker fixed to the back of M2. 
The intensity of the interference pattern is measured at a single point on the screen, indicated by the pink star in Fig~\ref{fig:interference_pattern}. 
The webcam records video in three colour channels; red, blue, and green. 
We use the green channel as an approximation of the total intensity produced by the monochromatic green laser.
The webcam samples at a rate of $30\, {\rm Hz}$, which limits the spectral content of observable signals to less than $15\,{\rm Hz}$, the Nyquist frequency.
